Type Error Telemetry at Scale: Pseudonymized and Stochiastic

### 2023-09-20 programming23/paper.tex
more text for more figures, 2 left for a first cut

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}

  % TODO use bars instead, percent only, don't care about counts
  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{} & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
    \mnonstrict{} & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
    \mstrict{} & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  \end{tabular}
  \caption{FILL error density}
  \label{f:error-delta}
\end{figure}

@@

@@

\subsection{Timeline}

\begin{figure}[t]\centering
  %% TODO can we 
  \mnonstrict{}\\
  \includegraphics[width=\columnwidth]{img/timeline-nonstrict.pdf}

@@

TODO correlations, compare modes!!!

>>>>

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes over time (seconds) to type error density (total errors / lines in project)}
  \label{f:error-density}
\end{figure}

\FILL{} what does the FS plot look like? is that a better nocheck-vs-rest comparison??

Moving on to type errors across the projects, and not just in the edit range,
a first interesting question is to study how the overall number of errors
changes over time.
But, the sheer number of errors can be skewed by codebase size.
To give a better basis for project-to-project comparisons, we focus
on the density of errors rather than the count, where \emph{density} is
the total number of type errors divided by the lines of code in the project.

\Cref{f:error-density} presents a zoomed-in view of type error density as it
changes over time in every session of the dataset.
The x-axis counts time in seconds from the start of the session.
The y-axis shows the change in density from the previous telemetry record
in the session to the next.
Changes can be positive or negative depending on whether the edits added
or removed errors.
The plots cut off at \pm{}50 density and 150 seconds.to focus on the
bulk of the data.
%% Sadly even with the zoom, it's too hard to follow one session ... the plot only gives a big-picture impression.

Due to the relative adoption of type analysis, there is much more
data for the weak \mnocheck{} mode than for \mnonstrict{},
and in turn for \mnonstrict{} relative to \mstrict{}.
Still, the plots reveal interesting points:

\begin{itemize}
  \item
    All three plots are roughly balanced,
    with equal ``mass'' above and below the x-axis.
    The next section explores this point in detail.

  \item
    After the 40-second mark, changes to type error density become much
    smaller.
    Why?
    There are also fewer sessions (\mstrict{} makes that very clear).

  \item
    \mnonstrict{} has the biggest number of up/down deltas in the 10+ range.
\end{itemize}

\subsection{Changes to Errors by Mode}

\begin{figure}[t]\centering
  % TODO use bars instead, percent only, don't care about counts
  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{} & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
    \mnonstrict{} & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
    \mstrict{} & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the total number of type errors?}
  \label{f:error-changes}
\end{figure}

\FILL{}

@@

\FILL{}

@@

\subsection{Session Timelines}

\begin{figure}[t]\centering
  %% TODO any way to cluster timelines to avoid the "creepiness" of spying on one user?
%  \mnonstrict{}\\
%  \includegraphics[width=\columnwidth]{img/timeline-nonstrict.pdf}

@@

Another insightful visualization that the telemetry supports is plots of individual sessions
over time.
\Cref{f:indy-session} shows the type errors, forced strict errors, and module switches
in two \mstrict{} mode sessions.
The x-axis shows time in seconds from the start of the session.
The y-axis counts errors.
Type errors appear as orange plus marks.
Forced strict errors appear as green x marks.
The vertical lines show module switches.

In the first plot, the error counts are low until about the 3,000s mark,
at which point an edit (not a module switch) leads to 15 errors.
These 15 errors disappear two records later.
Then, after two module switches and some edits, the number of errors spikes again.
In the second plot, the number of errors spikes in the beginning due to edits,
then remains low for the next few thousand seconds.

A general observation is that the number of type errors is always less than the number of forced strict
errors, though the numbers are close.
This should happen in \mstrict{} mode.
